### 2021-11-24 thesis/tex/nsiTheory.tex
update

We have no independent sensitivity for the neither chirality nor flavor type of $\epsilon^X$, so we sum over these and study the effective matter NSI parameter
 $\epsilon_{\ab}$:
\begin{align}
    \epsilon_{\ab} = \sum_{X \in \{L,R\}} \sum_{f \in \{e,u,d\}} \frac{N_f}{N_e} \epsilon^{fX}_{\ab}\,.
\end{align}
Our matter study will be wholly confined to the interior of the Earth, where we assume electrical neutrality and equal distribution of neutrons and protons, 
we get $N_u/N_e \simeq N_d/N_e \simeq 3$. Also we assume the components $\epsilon_{\a\b}$ to be real. Thus,
\begin{align}
    \epsilon_{\ab} =  \sum_X \epsilon_{\ab}^{eX} + 3(\epsilon_{\ab}^{uX} + \epsilon_{\ab}^{dX})
\end{align}
Now, $\epsilon_{\ab}$ enters the Hamiltonian as entries of a potential-like matrix. In Eq.~\ref{eq:NSIH}, $A_{CC}\text{diag}(1,0,0)$ is our 
familiar matter potential from the Standard Model. There is also our new term, $A_{CC} \epsilon$, which contains the components $\epsilon_{\a\b}$:
\begin{align}\label{eq:NSIH}
    H &= \frac{1}{2E} \left[UM^2U^\dagger + A_{CC}\,\text{diag}(1,0,0) + A_{CC}\, \epsilon \right] \nonumber \\
      &= \frac{1}{2E} \left[UM^2U^\dagger + A_{CC}
      \begin{pmatrix}
          1 + \epsilon_{ee} & \epsilon_{e\mu} & \epsilon_{e\tau}  \\
          \epsilon_{\mu e} & \epsilon_{\mu\mu} & \epsilon_{\mu\tau}  \\
          \epsilon_{\tau e} & \epsilon_{\tau\mu} & \epsilon_{\tau\tau}
      \end{pmatrix} \right]\,.
\end{align} 
In the limit $\epsilon_{\ab} \to 0$, we recover the standard interaction Hamiltonian from Eq.~\ref{eq:H_3gen}.
We can draw several conclusions from this form of the Hamiltonian. Any nonzero off-diagonal element $\epsilon_{\ab}, \a \neq \b$ will make the NSI violate 
lepton flavor, just as the off-diagonal elements of $U$ does in the SM. Moreover, since the SM potential has the same order in $A_{CC}$ as the NSIs, 
any $\epsilon_{\ab} \sim 1$ will make the new matter effect be the same order as the SM effect.

@@

Another feature of our DeepCore study includes the fact that we now have access to cascade events, in which $\ne$ and $\nt$ are more abundant. %TODO:show this?
Thus, we are no longer constrained to the $\mu$ channel alone, but we can now find interesting features in the other channels too. However, we 
remember that the $\nm$ flux is still the most abundant. %TODO: link to plot comparison of m and e flux.

@@

We plot the event pull $(N_{NSI} - N_{SI})/\sqrt{N_{SI}}$ where $N_{(N)SI}$ are the numbers of expected events
assuming (non-)standard interactions in Fig.~\ref{fig:event_pulls}. This gives the normalized difference in the
number of expected events at the detector, and illustrates the expected sensitivity of DeepCore for the NSI parameters.
%TODO: clean up above and expand a bit.
\begin{figure}
    \begin{center}
        \includegraphics[width=0.9\textwidth]{figures/Pee_eet_probs.pdf}
        \caption{Electron neutrino and antineutrino survival probabilities for
        $\ztrue = -1$ when $\eet = 0.2$. All other NSI parameters are fixed to zero. Compared to the $\nm$ and $\anm$ plots, here we see a strong difference across the whole energy range.}
        \label{fig:Pee_eet_probs}
    \end{center}
\end{figure}

>>>>

We have no independent sensitivity for the neither chirality nor flavor type of $\epsilon^X$, so we sum over these and study the effective matter NSI parameter
 $\epsilon_{\ab}$:
\begin{align}
    \epsilon_{\ab} = \sum_{X \in \{L,R\}} \sum_{f \in \{e,u,d\}} \frac{N_f}{N_e} \epsilon^{fX}_{\ab}\,.
\end{align}
Our matter study will be wholly confined to the interior of the Earth, where we assume electrical neutrality and equal distribution of neutrons and protons, 
we get $N_u/N_e \simeq N_d/N_e \simeq 3$. Also we assume the components $\epsilon_{\a\b}$ to be real. Thus,
\begin{align}
    \epsilon_{\ab} =  \sum_X \epsilon_{\ab}^{eX} + 3(\epsilon_{\ab}^{uX} + \epsilon_{\ab}^{dX})
\end{align}
Now, $\epsilon_{\ab}$ enters the Hamiltonian as entries of a potential-like matrix. In Eq.~\ref{eq:NSIH}, $A_{CC}\text{diag}(1,0,0)$ is our 
familiar matter potential from the Standard Model. There is also our new term, $A_{CC} \epsilon$, which contains the components $\epsilon_{\a\b}$:
\begin{align}\label{eq:NSIH}
    H &= \frac{1}{2E} \left[UM^2U^\dagger + A_{CC}\,\text{diag}(1,0,0) + A_{CC}\, \epsilon \right] \nonumber \\
      &= \frac{1}{2E} \left[UM^2U^\dagger + A_{CC}
      \begin{pmatrix}
          1 + \epsilon_{ee} & \epsilon_{e\mu} & \epsilon_{e\tau}  \\
          \epsilon_{\mu e} & \epsilon_{\mu\mu} & \epsilon_{\mu\tau}  \\
          \epsilon_{\tau e} & \epsilon_{\tau\mu} & \epsilon_{\tau\tau}
      \end{pmatrix} \right]\,.
\end{align} 
In the limit $\epsilon_{\ab} \to 0$, we recover the standard interaction Hamiltonian from Eq.~\ref{eq:H_3gen}.
We can draw several conclusions from this form of the Hamiltonian. Any nonzero off-diagonal element $\epsilon_{\ab}, \a \neq \b$ contribute
to neutrino mixing, just as the off-diagonal elements of $U$ does in the SM. Moreover, since the SM potential has the same order in $A_{CC}$ as the NSIs, 
any $\epsilon_{\ab} \sim 1$ will make the new matter effect be the same order as the SM effect.

@@

Another feature of our DeepCore study includes the fact that we now have access to cascade events, in which $\ne$ and $\nt$ are more abundant.
Thus, we are no longer constrained to the $\mu$ channel alone, but we can now find interesting features in the other channels too. However, we 
remember that the $\nm$ flux is still the most abundant.

@@

We plot the event pull $(N_{NSI} - N_{SI})/\sqrt{N_{SI}}$ where $N_{(N)SI}$ are the numbers of expected events
assuming (non-)standard interactions in Fig.~\ref{fig:event_pulls}. This gives the normalized difference in the
number of expected events at the detector, and illustrates the expected sensitivity of DeepCore for the NSI parameters.

\begin{figure}
    \begin{center}
        \includegraphics[width=0.9\textwidth]{figures/Pee_eet_probs.pdf}
        \caption{Electron neutrino and antineutrino survival probabilities for
        $\ztrue = -1$ when $\eet = 0.2$. All other NSI parameters are fixed to zero. Compared to the $\nm$ and $\anm$ plots, here we see a strong difference across the whole energy range.}
        \label{fig:Pee_eet_probs}
    \end{center}
\end{figure}

### 2021-11-24 thesis/tex/oscillations.tex

We are now down to six degrees of freedom in the three dimensional case: three angles
of the form $s_{ij}, c_{ij}$, and three phase of the form $e^{i\delta_\mathrm{CP}}$ and $e^{\lambda{i}}$. The angle $\theta_{ij}$ specifies the amount of rotation in the $i-j$ plane, allowing an $i$ eigenstate to mix with a $j$ eigenstate. The physical interpretation of
a mixing angle is then the degree to which the $\nu_i$ mass eigenstate mixes with the $\nu_j$ mass eigenstate. Thus we refer to $\theta_{ij}$ as the mixing angles. A non-zero mixing angle will then 
cause the flavor eigenstate $\nu_\alpha$ to be a superposition of the mass eigenstates $\nu_i$ that we allowed to mix. More on this in Section~\ref{sec:prop}.

@@

From the expression for the weak charged current in~\ref{eq:j_CC3}, we see that the summation over the mass index $k$ with the mixing matrix elements constructs a flavor neutrino, which interacts with the charged lepton field $\ell_{\a L}$.
In other words, the charged current generates a flavor neutrino $\na$, which is a superposition of the mass eigenstates 
$\nu_k$ with weights $U_{\a k}^\dagger$. In the ket-formalism, we express this as
\begin{align}\label{eq:osc_1}
    \ket{\na} = \sum_k U^\dagger_{\a k} \ket{\nu_k}\,.
\end{align}
It is the mass eigenstates $\ket{\nu_k}$ that are eigenstates of the Hamiltonian, with eigenvalues
\begin{align}\label{eq:disp}
    E_k = \sqrt{\vec{p}^2 + m_k^2}\,.
\end{align}
The solution to the time-dependent Schrödinger equation 
\begin{align}\label{eq:TDSE}
    i \dv{}{t} \ket{\nu_k(t)} = H_0\ket{\nu_k(t)}\,,
\end{align}
where $H_0$ is the vacuum Hamiltonian. 
The solution to Eq.~\ref{eq:TDSE} gives ut the time evolution
in the form of plane wave solutions: 
\begin{align}
    \ket{\nu_k(t)} = e^{-iE_kt}\ket{\nu_k}\,.
\end{align}
Inserting the plane wave solution into Eq.~\ref{eq:osc_1}, we get 
\begin{align}\label{eq:nat}
    \ket{\na(t)} = \sum_k U^\dagger_{\a k} e^{-iE_kt}\ket{\nu_k}\,.
\end{align}
Now we know how to evolve and combine the mass eigenstates to form a flavor eigenstate, but how about the reverse? We swap the index $k\to j$ in Eq.~\ref{eq:osc_1} and multiply by $U_{\a k}$:
\begin{align}\label{eq:osc_k}
    \sum_\a U_{\a k}\ket{\na} &= \sum_{\alpha, j} U_{\a k} U_{\a j}^\dagger \ket{\nu_j} \nonumber \\
                              &= \sum_{j} \delta_{kj} \ket{\nu_j} \nonumber \\
                              &= \ket{\nu_k}\,,
\end{align}
where we have used the unitarity of the leptonic mixing matrix. Eqs.~\ref{eq:osc_1} and \ref{eq:nat} yield 
\begin{align}
    \ket{\na(t)} &= \sum_k U^\dagger_{\a k} e^{-iE_kt}\ket{\nu_k} \nonumber \\
                 &= \sum_k U^\dagger_{\a k} e^{-iE_kt}\left(\sum_\b U_{\b k}\ket{\nb}\right) \nonumber \\
                 &= \sum_{k,\b} U^\dagger_{\a k}U_{\b k} e^{-iE_kt} \ket{\nb}\,.
\end{align}
The probability of the flavor transition $\na \to \nb$ at time $t$ is $\abs{\bra{\nb}\ket{\na (t)}}^2$:
\begin{align}
    P_{\na \to \nb}(t) &= \sum_{k,j} \abs{{\bra{\nb}\ket{\na (t)}}}^2 \nonumber \\
    &= U^\dagger_{\a k}U_{\b k}U^\dagger_{\b j}U_{\a j} e^{-i(E_k-E_j)t}\,.
\end{align}

We assume the neutrino masses $m_k$ to be extremely small compared to their associated energies $E_k$. 
Thus, $v\approx 1$, and $\abs{\vec{p}} \approx E$ making the energy-dispersion relation of Eq.~\ref{eq:disp} to first order:
\begin{align}\label{eq:ultra_rel}
    E_k &= \sqrt{{\vec{p}}^2 + m_k^2} \nonumber \\
        &= {\vec{p}}^2\sqrt{1 + \frac{m_k^2}{\vec{p}^2}} \nonumber \\
        &\approx E + \frac{m_k^2}{2E}
\end{align}
Hence, the exponential can be simplified, and simplifying the notation $P_{\na \to \nb}(t) \to P_{\ab}(t)$ we get 
\begin{align}
    P_{\ab}(t) = \sum_{k,j} \abs{U^\dagger_{\a k}U_{\b k}U^\dagger_{\b j}U_{\a j}} e^{-i(m_k^2-m_j^2)t/2E}\,.
\end{align}
Now, our approximation $v\approx 1$ implies $x\approx t $, thus
\begin{align}
    P_{\ab}(x) &= \sum_{k,j} \abs{U^\dagger_{\a k}U_{\b k}U^\dagger_{\b j}U_{\a j}} e^{-i(m_k^2-m_j^2)x/2E} \nonumber \\
                       &= \sum_{k,j} \abs{U^\dagger_{\a k}U_{\b k}U^\dagger_{\b j}U_{\a j}} \exp(-i\frac{\dm_{kj}x}{2E})\,,
\end{align}
where we in the last step have defined the \emph{mass-squared difference} $\dm_{kj} = m_k^2 - m_j^2$. Since the oscillation probability depends on this quantity
rather than the individual masses, it is impossible to measure the absolute mass $m_k$ through neutrino oscillations. We have
to rely on other types of experiments to get us at least one absolute mass, from which we then can calculate the remaining.
Squaring the unitarity condition $\sum_k U_{\a k}U_{\b k}^\dagger = \delta_{\a \b}$ yields 
\begin{align}
    \sum_k \abs{U_{\a k}}^2\abs{U_{\b k}}^2 = \delta_{\a \b} - 2 \sum_{k>j}\Re[U_{\a k}^\dagger U_{\b k} U_{\a j} U_{\b j}^\dagger]
\end{align}
We have
\begin{align}\label{eq:Pab} 
    P_{\ab} &= \sum_{k,j} \abs{U_{\a k} ^ \dagger U_{\b k}U_{\b j}^\dagger U_{\a j}} \exp(-i\frac{\dm_{kj}x}{2E}) \nonumber \\
            &=  \sum_k \abs{U_{\a k}}^2\abs{U_{\b k}}^2 + \sum_{k \neq j} \abs{U_{\a k} ^ \dagger U_{\b k}U_{\b j}^\dagger U_{\a j}} \exp(-i\frac{\dm_{kj}x}{2E}) \nonumber \\
            &= \delta_{\a \b} - 2 \sum_{k>j}\Re[U_{\a k}^\dagger U_{\b k} U_{\a j} U_{\b j}^\dagger] +\sum_{k \neq j} \abs{U_{\a k} ^ \dagger U_{\b k}U_{\b j}^\dagger U_{\a j}} \exp(-i\frac{\dm_{kj}x}{2E}) \nonumber \\
            %&= \delta_{\a \b} - 2 \sum_{k>j}\Re[U_{\a k}^\dagger U_{\b k} U_{\a j} U_{\b j}^\dagger] + 2\sum_{k > j} U_{\a k}U_{\b k}U_{\b j}U_{\a j} \exp(-i\frac{\dm_{kj}x}{2E}) \nonumber \\
            %&= \delta_{\a \b} - 2\sum_{k > j} U_{\a k}U_{\b k}U_{\b j}U_{\a j} \left[1-\exp(-i\frac{\dm_{kj}x}{2E})\right] \nonumber \\
            %&= \delta_{\a \b} - 2\sum_{k > j} U_{\a k}U_{\b k}U_{\b j}U_{\a j} \left[1-\cos(\frac{\dm_{kj}x}{2E})\right] \nonumber \\
            &= \delta_{\a \b} - 4\sum_{k > j} \Re[U_{\a k} ^ \dagger U_{\b k}U_{\b j}^\dagger U_{\a j}] \sin^2{\left(\frac{\dm_{kj}x}{4E}\right)} \nonumber \\
            &\hspace{1.2cm}+ 2\sum_{k > j} \Im[U_{\a k} ^ \dagger U_{\b k}U_{\b j}^\dagger U_{\a j}] \sin{\left(\frac{\dm_{kj}x}{4E}\right)}\,,
\end{align}
which is the probability of neutrino vacuum oscillations.
A similar calculation for antineutrinos yield  
\begin{align}\label{eq:Pab} 
    P_{\ab} 
            &= \delta_{\a \b} - 4\sum_{k > j} \Re[U_{\a k} ^ \dagger U_{\b k}U_{\b j}^\dagger U_{\a j}] \sin^2{\left(\frac{\dm_{kj}x}{4E}\right)} \nonumber \\
            &\hspace{1.2cm}- 2\sum_{k > j} \Im[U_{\a k} ^ \dagger U_{\b k}U_{\b j}^\dagger U_{\a j}] \sin{\left(\frac{\dm_{kj}x}{4E}\right)}\,,
\end{align}

Experimental measurements show that $\theta_{13}$ is the smallest of the mixing angles~\cite{nufit}, causing 
$\nm$ to primarily mix into $\nt$. Using values from~\cite{nufit}, but taking $\delta_\mathrm{CP} = 0$ for simplicity,
the mixing matrix from~\ref{PMNS_def} takes the values
\begin{align}\label{eq:Uvalues}
   U = \begin{pmatrix}
       U_{e 1} & U_{e2} & U_{e3} \\
       U_{\mu 1} & U_{\mu 2} & U_{\mu 3} \\
       U_{\tau 1} & U_{\tau 2} & U_{\tau 3}
   \end{pmatrix} 
   = \begin{pmatrix}
       0.825 & 0.545 & 0.149 \\
       -0.455 & 0.485 & 0.746 \\
       0.334 & -0.684 & 0.649
   \end{pmatrix} \,,
\end{align}
which use to we plot the $P_{\mu \beta}$ probabilities in Fig.~\ref{fig:vac_osc}. Here, we show the energy spectra of these oscillations in the low \si{\GeV} range for neutrinos that travel \SI{12000}{\km}\footnote{This baseline
is approximately the Earth diameter, and will be the maximum travel distance which we will use later on.}.
In the left-most panel, we see a suppressed $\nm \to \ne$ transition. We can understand this behavior by studying Eq.~\ref{eq:Pab} and the 
numerical values of the mixing matrix in Eq.~\ref{eq:Uvalues}.
The smallness of $U_{e3}$ reduces the effect of $\dm[21]$ mass-squared difference,
allowing $\dm[31]$ to drive the oscillations. 
The most important mixing matrix 
elements for $P_{\a\b}$ are $U_{\a3}, U_{\a1}$ and $U_{\b3}, U_{\b1}$. %TODO: write whole 4 product here
%instead of U,U and U,U

\begin{figure}[!h]
    %TODO: write pme, pmm, pmt in titles, bigger ticks
    \begin{center}
    \includegraphics[width=1\textwidth]{figures/vac_osc.pdf}
    \caption{Probabilities of $\nm$ oscillations travelling 12 000 km in vaccum.}\label{fig:vac_osc}
    \end{center}
\end{figure}
This form elucidates an important aspect of neutrino oscillations. The mixing matrix elements
determine the amplitude of the oscillations, while the mass-squared differences together with the ratio $L/E$ determine the
frequency. So in vacuum, the mixing angles only influence the oscillation amplitude, while the energy, trajectory, and masses of
the neutrinos only influence the oscillation frequency.

>>>>

We are now down to six degrees of freedom in the three dimensional case: three angles
of the form $s_{ij}, c_{ij}$, and three phase of the form $e^{i\delta_\mathrm{CP}}$ and $e^{i\lambda_{i}}$. The angle $\theta_{ij}$ specifies the amount of rotation in the $i-j$ plane, allowing an $i$ eigenstate to mix with a $j$ eigenstate. The physical interpretation of
a mixing angle is then the degree to which the $\nu_i$ mass eigenstate mixes with the $\nu_j$ mass eigenstate. Thus we refer to $\theta_{ij}$ as the mixing angles. A non-zero mixing angle will then 
cause the flavor eigenstate $\nu_\alpha$ to be a superposition of the mass eigenstates $\nu_i$ that we allowed to mix. More on this in Section~\ref{sec:prop}.

@@

From the expression for the weak charged current in~\ref{eq:j_CC3}, we see that the summation over the mass index $k$ with the mixing matrix elements constructs a flavor neutrino, which interacts with the charged lepton field $\ell_{\a L}$.
In other words, the charged current generates a flavor neutrino $\na$, which is a superposition of the mass eigenstates 
$\nu_k$ with weights $U_{\a k}^*$. In the ket-formalism, we express this as
\begin{align}\label{eq:osc_1}
    \ket{\na} = \sum_k U^*_{\a k} \ket{\nu_k}\,.
\end{align}
It is the mass eigenstates $\ket{\nu_k}$ that are eigenstates of the Hamiltonian, with eigenvalues
\begin{align}\label{eq:disp}
    E_k = \sqrt{\vec{p}^2 + m_k^2}\,.
\end{align}
The solution to the time-dependent Schrödinger equation 
\begin{align}\label{eq:TDSE}
    i \dv{}{t} \ket{\nu_k(t)} = H_0\ket{\nu_k(t)}\,,
\end{align}
where $H_0$ is the vacuum Hamiltonian. 
The solution to Eq.~\ref{eq:TDSE} gives ut the time evolution
in the form of plane wave solutions: 
\begin{align}
    \ket{\nu_k(t)} = e^{-iE_kt}\ket{\nu_k}\,.
\end{align}
Inserting the plane wave solution into Eq.~\ref{eq:osc_1}, we get 
\begin{align}\label{eq:nat}
    \ket{\na(t)} = \sum_k U^*_{\a k} e^{-iE_kt}\ket{\nu_k}\,.
\end{align}
Now we know how to evolve and combine the mass eigenstates to form a flavor eigenstate, but how about the reverse? We swap the index $k\to j$ in Eq.~\ref{eq:osc_1} and multiply by $U_{\a k}$:
\begin{align}\label{eq:osc_k}
    \sum_\a U_{\a k}\ket{\na} &= \sum_{\alpha, j} U_{\a k} U_{\a j}^* \ket{\nu_j} \nonumber \\
                              &= \sum_{j} \delta_{kj} \ket{\nu_j} \nonumber \\
                              &= \ket{\nu_k}\,,
\end{align}
where we have used the unitarity of the leptonic mixing matrix. Eqs.~\ref{eq:osc_1} and \ref{eq:nat} yield 
\begin{align}
    \ket{\na(t)} &= \sum_k U^*_{\a k} e^{-iE_kt}\ket{\nu_k} \nonumber \\
                 &= \sum_k U^*_{\a k} e^{-iE_kt}\left(\sum_\b U_{\b k}\ket{\nb}\right) \nonumber \\
                 &= \sum_{k,\b} U^*_{\a k}U_{\b k} e^{-iE_kt} \ket{\nb}\,.
\end{align}
The probability of the flavor transition $\na \to \nb$ at time $t$ is $\abs{\bra{\nb}\ket{\na (t)}}^2$:
\begin{align}
    P_{\na \to \nb}(t) &= \sum_{k,j} \abs{{\bra{\nb}\ket{\na (t)}}}^2 \nonumber \\
    &= U^*_{\a k}U_{\b k}U^*_{\b j}U_{\a j} e^{-i(E_k-E_j)t}\,.
\end{align}

We assume the neutrino masses $m_k$ to be extremely small compared to their associated energies $E_k$. 
Thus, $v\approx 1$, and $\abs{\vec{p}} \approx E$ making the energy-dispersion relation of Eq.~\ref{eq:disp} to first order:
\begin{align}\label{eq:ultra_rel}
    E_k &= \sqrt{{\vec{p}}^2 + m_k^2} \nonumber \\
        &= {\vec{p}}^2\sqrt{1 + \frac{m_k^2}{\vec{p}^2}} \nonumber \\
        &\approx E + \frac{m_k^2}{2E}
\end{align}
Hence, the exponential can be simplified, and simplifying the notation $P_{\na \to \nb}(t) \to P_{\ab}(t)$ we get 
\begin{align}
    P_{\ab}(t) = \sum_{k,j} \abs{U^*_{\a k}U_{\b k}U^*_{\b j}U_{\a j}} e^{-i(m_k^2-m_j^2)t/2E}\,.
\end{align}
Now, our approximation $v\approx 1$ implies $x\approx t $, thus
\begin{align}
    P_{\ab}(x) &= \sum_{k,j} \abs{U^*_{\a k}U_{\b k}U^*_{\b j}U_{\a j}} e^{-i(m_k^2-m_j^2)x/2E} \nonumber \\
                       &= \sum_{k,j} \abs{U^*_{\a k}U_{\b k}U^*_{\b j}U_{\a j}} \exp(-i\frac{\dm_{kj}x}{2E})\,,
\end{align}
where we in the last step have defined the \emph{mass-squared difference} $\dm_{kj} = m_k^2 - m_j^2$. Since the oscillation probability depends on this quantity
rather than the individual masses, it is impossible to measure the absolute mass $m_k$ through neutrino oscillations. We have
to rely on other types of experiments to get us at least one absolute mass, from which we then can calculate the remaining.
Squaring the unitarity condition $\sum_k U_{\a k}U_{\b k}^* = \delta_{\a \b}$ yields 
\begin{align}
    \sum_k \abs{U_{\a k}}^2\abs{U_{\b k}}^2 = \delta_{\a \b} - 2 \sum_{k>j}\Re[U_{\a k}^* U_{\b k} U_{\a j} U_{\b j}^*]
\end{align}
We have
\begin{align}\label{eq:Pab} 
    P_{\ab} &= \sum_{k,j} \abs{U_{\a k} ^ * U_{\b k}U_{\b j}^* U_{\a j}} \exp(-i\frac{\dm_{kj}x}{2E}) \nonumber \\
            &=  \sum_k \abs{U_{\a k}}^2\abs{U_{\b k}}^2 + \sum_{k \neq j} \abs{U_{\a k} ^ * U_{\b k}U_{\b j}^* U_{\a j}} \exp(-i\frac{\dm_{kj}x}{2E}) \nonumber \\
            &= \delta_{\a \b} - 2 \sum_{k>j}\Re[U_{\a k}^* U_{\b k} U_{\a j} U_{\b j}^*] +\sum_{k \neq j} \abs{U_{\a k} ^ * U_{\b k}U_{\b j}^* U_{\a j}} \exp(-i\frac{\dm_{kj}x}{2E}) \nonumber \\
            %&= \delta_{\a \b} - 2 \sum_{k>j}\Re[U_{\a k}^* U_{\b k} U_{\a j} U_{\b j}^*] + 2\sum_{k > j} U_{\a k}U_{\b k}U_{\b j}U_{\a j} \exp(-i\frac{\dm_{kj}x}{2E}) \nonumber \\
            %&= \delta_{\a \b} - 2\sum_{k > j} U_{\a k}U_{\b k}U_{\b j}U_{\a j} \left[1-\exp(-i\frac{\dm_{kj}x}{2E})\right] \nonumber \\
            %&= \delta_{\a \b} - 2\sum_{k > j} U_{\a k}U_{\b k}U_{\b j}U_{\a j} \left[1-\cos(\frac{\dm_{kj}x}{2E})\right] \nonumber \\
            &= \delta_{\a \b} - 4\sum_{k > j} \Re[U_{\a k} ^ * U_{\b k}U_{\b j}^* U_{\a j}] \sin^2{\left(\frac{\dm_{kj}x}{4E}\right)} \nonumber \\
            &\hspace{1.2cm}+ 2\sum_{k > j} \Im[U_{\a k} ^ * U_{\b k}U_{\b j}^* U_{\a j}] \sin{\left(\frac{\dm_{kj}x}{4E}\right)}\,,
\end{align}
which is the probability of neutrino vacuum oscillations.
A similar calculation for antineutrinos yield  
\begin{align}
    P_{\ab} 
            &= \delta_{\a \b} - 4\sum_{k > j} \Re[U_{\a k} ^ * U_{\b k}U_{\b j}^* U_{\a j}] \sin^2{\left(\frac{\dm_{kj}x}{4E}\right)} \nonumber \\
            &\hspace{1.2cm}- 2\sum_{k > j} \Im[U_{\a k} ^ * U_{\b k}U_{\b j}^* U_{\a j}] \sin{\left(\frac{\dm_{kj}x}{4E}\right)}\,.
\end{align}

This form elucidates an important aspect of neutrino oscillations. The mixing matrix elements
determine the amplitude of the oscillations, while the mass-squared differences together with the ratio $L/E$ determine the
frequency. So in vacuum, the mixing angles only influence the oscillation amplitude, while the energy, trajectory, and masses of
the neutrinos only influence the oscillation frequency.

Experimental measurements show that $\theta_{13}$ is the smallest of the mixing angles~\cite{nufit}, causing 
$\nm$ to primarily mix into $\nt$. Using values from~\cite{nufit}, but taking $\delta_\mathrm{CP} = 0$ for simplicity,
the mixing matrix from~\ref{PMNS_def} takes the values
\begin{align}\label{eq:Uvalues}
   U = \begin{pmatrix}
       U_{e 1} & U_{e2} & U_{e3} \\
       U_{\mu 1} & U_{\mu 2} & U_{\mu 3} \\
       U_{\tau 1} & U_{\tau 2} & U_{\tau 3}
   \end{pmatrix} 
   = \begin{pmatrix}
       0.825 & 0.545 & 0.149 \\
       -0.455 & 0.485 & 0.746 \\
       0.334 & -0.684 & 0.649
   \end{pmatrix} \,,
\end{align}
which use to we plot the $P_{\mu \beta}$ probabilities in Fig.~\ref{fig:vac_osc}. Here, we show the energy spectra of these oscillations in the low \si{\GeV} range for neutrinos that travel \SI{12000}{\km}\footnote{This baseline
is approximately the Earth diameter, and will be the maximum travel distance which we will use later on.}.
In the left-most panel, we see a suppressed $\nm \to \ne$ transition. We can understand this behavior by studying Eq.~\ref{eq:Pab} and the 
numerical values of the mixing matrix in Eq.~\ref{eq:Uvalues}.
The smallness of $U_{e3}$ reduces the effect of $\dm[21]$ mass-squared difference,
allowing $\dm[31]$ to drive the oscillations. Thus, 
The most important mixing matrix elements for $P_{\a\b}$ are the ones appearing in the term with $\dm[31]$:
namely the product $U_{\a3}^*U_{\b3} U_{\b1}^*U_{\a1}$.

\begin{figure}[!h]
    \begin{center}
    \includegraphics[width=1\textwidth]{figures/vac_osc.pdf}
    \caption{Probabilities of $\nm$ oscillations after travelling 12 000 km in vaccum.}\label{fig:vac_osc}
    \end{center}
\end{figure}
